One-arm spirals in self-gravitating dead zones in protoplanetary discs

### 2014-10-21 paper/model.tex
minor update

Our fiducial parameter values are: $s=2$, $R_{d1}=R_0$, $R_{d2}=2R_0$,
$\epsilon=0.1$, $\delta R_d=5$, $h=0.05$ and
$Q_\mathrm{out}=2$. Initially there is no vertical or radial velocity
($v_R = v_r = v_\theta = 0$). The azimuthal velocity is initialized to
satisfy centrifugal balance with pressure and gravity,
\begin{align}
  \frac{v_\phi^2}{r} = \frac{1}{\rho}\frac{\p p}{\p r} + \frac{\p
    \Phi_\mathrm{tot}}{\p r}
\end{align}
and similarly in 2D. The angular velocity is $\Omega = v_\phi/R$.

@@

\subsection{FARGO}
Our primary code is FARGO with self-gravity \citep{baruteau08}. This
is a popular, simple finite-difference code for 2D discs. `FARGO' refers
to its azimuthal transport algorithm, which circumvents time-step
contraint set by the inner disc boundary. 
The 2D Poisson equation is solved in integral form 
\begin{align}\label{2d_grav}
  \Phi_{d,z=0}(R,\phi) = - \int
  \frac{G\Sigma(R^\prime,\phi^\prime)}{\sqrt{R^2+R^{\prime 2} -
      2RR^\prime\cos{(\phi - \phi^\prime)} + \epsilon_g^2}} 
\end{align}
using Fast Fouier Transform (FFT), where $\epsilon_g$ is a softening
length to prevent a numerical singularity and the integration is taken
over the disc. The FFT approach requires
$\epsilon_g\propto R$ \citep{baruteau08}.

@@

The non-axisymmetric surface density can be visualised with
\begin{align}
  \Delta \Sigma = \frac{\Sigma - \Sigma_{00}}{\Sigma_{00}}, 
\end{align}
where $\Sigma_{00} = \Sigma_0(t=0)$, while an individual mode as
\begin{align}
  \Delta\Sigma_m = \frac{2}{\Sigma_{00}} \real\left[\Sigma_m \exp{(\ii
      m\phi)}\right].
\end{align}

@@

\subsection{Linear density waves}
We will also use linear density wave theory to interpret
simulation data. Here, we follow conventions in \cite{shu91} and quote
results therein. The dispersion relation for tightly-wound density
waves of the form 
$\exp{[\ii(\omega t - m \phi + kR)]}$ in a razor-thin
isothermal disc is 
\begin{align}\label{dispersion}
  (\omega - m\Omega)^2 = \kappa^2 + k^2c_s^2 - 2\pi G \Sigma |k|, 
\end{align}
where $\kappa^2 = R^{-3}\p_R(R^4\Omega^2)$ is the square of the epicycle frequency, 
and $k$ is a real wavenumber such that $|kR|\gg1$.

>>>>

Our fiducial parameter values are: $s=2$, $R_{d1}=R_0$, $R_{d2}=2R_0$,
$\epsilon=0.1$, $\delta R_d=5$, $h=0.05$ and
$Q_\mathrm{out}=2$. An example of the initial surface density and the
Toomre parameter (see\ref{wkb}) is shown in Fig. \ref{initial_surf}. 
Initially there is no vertical or radial velocity
($v_R = v_r = v_\theta = 0$). The azimuthal velocity is initialized to
satisfy centrifugal balance with pressure and gravity,
\begin{align}
  \frac{v_\phi^2}{r} = \frac{1}{\rho}\frac{\p p}{\p r} + \frac{\p
    \Phi_\mathrm{tot}}{\p r}
\end{align}
and similarly in 2D. The angular velocity is $\Omega = v_\phi/R$.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.7cm 0cm
  0cm]{figures/compare_profiles_dens000} 
  \includegraphics[width=\linewidth]{figures/compare_profiles_Q000}
  \caption{Fiducial profiles of the surface density (top) and Toomre
    parameter (bottom) used in this work.\label{initial_surf}}
\end{figure}

@@

\subsection{FARGO}
Our primary code is FARGO with self-gravity \citep{baruteau08}. This
is a popular, simple finite-difference code for 2D discs. `FARGO' refers
to its azimuthal transport algorithm, which circumvents time-step
contraint set by the inner disc boundary. 
The 2D Poisson equation is solved in integral form 
\begin{align}\label{2d_grav}
  \Phi_{d,z=0}(R,\phi) = - \int
  \frac{G\Sigma(R^\prime,\phi^\prime)R^\prime dR^\prime d\phi^\prime}{\sqrt{R^2+R^{\prime 2} -
      2RR^\prime\cos{(\phi - \phi^\prime)} + \epsilon_g^2}} 
\end{align}
using Fast Fouier Transform (FFT), where $\epsilon_g$ is a softening
length to prevent a numerical singularity and the integration is taken
over the disc. The FFT approach requires
$\epsilon_g\propto R$ \citep{baruteau08}.

@@

The non-axisymmetric surface density can be visualised with
\begin{align}
  \Delta \Sigma = \frac{\Sigma - \Sigma_{0}}{\Sigma_{0}}, 
\end{align}
while an individual mode as
\begin{align}
  \Delta\Sigma_m = \frac{2}{\Sigma_{00}} \real\left[\Sigma_m \exp{(\ii
      m\phi)}\right]
	\end{align}
where $\Sigma_{00} = \Sigma_0(t=0)$.

@@

\subsection{Linear density waves}\label{wkb}
We will also use linear density wave theory to interpret
simulation data. Here, we follow conventions in \cite{shu91} and quote
results therein. The dispersion relation for tightly-wound density
waves of the form 
$\exp{[\ii(\omega t - m \phi + kR)]}$ in a razor-thin
isothermal disc is 
\begin{align}\label{dispersion}
  (\omega - m\Omega)^2 = \kappa^2 + k^2c_s^2 - 2\pi G \Sigma |k|, 
\end{align}
where $\kappa^2 = R^{-3}\p_R(R^4\Omega^2)$ is the square of the epicycle frequency, 
and $k$ is a real wavenumber such that $|kR|\gg1$.
